---
title: "Shetty_S_M8_SVM"
author: "Sooraj Shetty"
date: "March 20, 2017"
output: word_document
---

```{r}
library(kernlab)
library(e1071)
library(ggplot2)
library(openxlsx)
```

Loading my data - Abalone Data Set 

```{r}
data_url <- 'http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data'
ABA <- read.csv(url(data_url),sep=",",header = FALSE)
colnames(ABA) <- c("Sex","Length","Diameter","Height","WholeWeight","ShuckedWeight","VisceraWeight","ShellWeight","Rings")
head(ABA)
summary(ABA)
```

```{r}
ABAsvm <- ABA[c("Sex","Length","Diameter","Height","WholeWeight","ShuckedWeight","VisceraWeight","ShellWeight","Rings")]
ABAsvm$Sex <- as.factor(ABAsvm$Sex)
head(ABAsvm)
str(ABAsvm)
```

Divide Training Data into x (containing all features) and y (containing classes)

```{r}
x <- ABAsvm[c("Sex","Length","Diameter","Height","WholeWeight","ShuckedWeight","VisceraWeight","ShellWeight","Rings")]
y <- ABAsvm$Sex
```

Creating SVM Models and summary of each model

```{r}
svm_model_rad <- svm(ABAsvm$Sex ~ ., data=ABAsvm, kernel="radial")
summary(svm_model_rad)

svm_model_lin <- svm(ABAsvm$Sex ~ ., data=ABAsvm, kernel="linear")
summary(svm_model_lin)

svm_model_sig <- svm(ABAsvm$Sex ~ ., data=ABAsvm, kernel="sigmoid")
summary(svm_model_sig)

svm_model_pol <- svm(ABAsvm$Sex ~ ., data=ABAsvm, kernel="polynomial")
summary(svm_model_pol)
```

See the confusion matrix result of prediction, using command table to compare the result of SVM prediction and the class data in y variable.

```{r}
predR <- predict(svm_model_rad,x)
table(predR,y)
a1 <- predR == y
table(a1)

predL <- predict(svm_model_lin,x)
table(predL,y)
a2 <- predL == y
table(a2)

predS <- predict(svm_model_sig,x)
table(predS,y)
a3 <- predS == y
table(a3)

predP <- predict(svm_model_pol,x)
table(predP,y)
a4 <- predP == y
table(a4)
```

We can see from all the above confusion matrices that svm_model_rad is showing most correct predictions, so we will further tune svm_model_rad to find best cost and gamma

Tuning SVM to find the best cost and gamma.

```{}
# I am facing the following error when I run this section of the code

#NAs introduced by coercion
#Error in if (any(co)) { : missing value where TRUE/FALSE needed

# Could not rectify this chunk of the code.

svm_tune <- tune(svm, train.x=x, train.y=y, kernel="radial", ranges=list(cost=c(0.01,0.1,1,2,5,10), gamma=c(0.1,0.2,0.5,1,1.5,2)))
print(svm_tune)
```

You can create svm model again and try to run again

```{r}

svm_model_after_tune <- svm(ABAsvm$Sex ~ ., data=ABAsvm, kernel="radial", cost=1, gamma=0.125)

summary(svm_model_after_tune)
```

Run Prediction again with new model
See the confusion matrix result of prediction, using command table to compare the result of SVM prediction and the class data in y variable.

```{r}
pred <- predict(svm_model_after_tune,x)
table(pred,y)
a5 <- pred == y
table(a5)
```